---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
[![DOI](https://zenodo.org/badge/357362984.svg)](https://zenodo.org/badge/latestdoi/357362984)

# An Introductory Tutorial on Cohort State-Transition Models in R Using a Cost-Effectiveness Analysis Example

This GitHub repository provides the code of the tutorial on how to implement time-independent cohort state-transition models (cSTMs) in R using a cost-effectiveness analysis (CEA) example, explained in the following manuscript: 

- Alarid-Escudero F, Krijkamp EM, Enns EA, Yang A, Hunink MGM, Pechlivanoglou P, Jalal H. [An Introductory Tutorial on Cohort State-Transition Models in R Using a Cost-Effectiveness Analysis Example](http://arxiv.org/abs/2001.07824). arXiv:200107824v4. 2022:1-30.

The release that accompanies the published article has been archived in zenodo: https://zenodo.org/badge/latestdoi/357362984

The [`R`](https://github.com/DARTH-git/cohort-modeling-tutorial-intro/tree/main/R) folder includes two different scripts corresponding to functions used to synthesize cSTMs outputs and conduct several sensitivity analyses:

   - [`Funtions.R`](https://github.com/DARTH-git/cohort-modeling-tutorial-intro/blob/main/R/Functions.R): Functions that generate epidemiological measures from time-independent cSTMs and compute within-cycle correction, parameter transformation, matrix checks, and CEA and PSA visualization.
   - [`Functions_cSTM_time_indep.R`](https://github.com/DARTH-git/cohort-modeling-tutorial-intro/blob/main/R/Functions_cSTM_time_indep.R): These functions wrap the time-independent cSTM, compute CEA measures, and generate probabilistic sensitivity analysis (PSA) input datasets.

## How to cite this R code in your article

You can cite the R code in this repository like this "we based our analysis using the R code from Alarid-Escudero F et al. (2021)". Here is the full bibliographic reference to include in your reference list for the manuscript and the R code (don't forget to update the 'last accessed' date):

> Alarid-Escudero F, Krijkamp EM, Enns EA, Yang A, Hunink MGM, Pechlivanoglou P, Jalal H. An Introductory Tutorial on Cohort State-Transition Models in R Using a Cost-Effectiveness Analysis Example (http://arxiv.org/abs/2001.07824). arXiv:200107824v4. 2022:1-30. 

> Alarid-Escudero F, Krijkamp EM, Enns EA, Yang A, Hunink MGM, Pechlivanoglou P, Jalal H (2022). R Code for An Introductory Tutorial on Cohort State-Transition Models in R Using a Cost-Effectiveness Analysis Example (Version v0.2.1). Zenodo. [10.5281/zenodo.5223093](https://www.doi.org/10.5281/zenodo.5223093). Last accessed 30 March 2022.

If you adapted the code, you should indicate "Adapted from:" or "Based on" so it is understood that you modified the code. For more information on how to cite computer code, we refer the user to review [Writing Code (from MIT Research Guide)](https://integrity.mit.edu/handbook/writing-code), which provides examples of how and when to cite computer code.

## Preliminaries
- Install [RStudio](https://www.rstudio.com/products/rstudio/download/)
- Install [`dampack`](https://cran.r-project.org/web/packages/dampack/index.html) R package from CRAN
```{r, eval=FALSE}
# Install release version from CRAN
install.packages("dampack")

# Or install development version from GitHub
# devtools::install_github("DARTH-git/dampack")
```
- Install `devtools` to install [`darthtools`](https://github.com/DARTH-git/darthtools) R package from [DARTH's GitHub](https://github.com/DARTH-git)
```{r, eval=FALSE}
# Install release version from CRAN
install.packages("devtools")

# Or install development version from GitHub
# devtools::install_github("r-lib/devtools")
```
- Install `darthtools` using `devtools`
```{r, eval=FALSE}
# Install development version from GitHub
devtools::install_github("DARTH-git/darthtools")
```
We recommend familiarizing with the [DARTH](http://darthworkgroup.com) coding framework described in 

- Alarid-Escudero F, Krijkamp EM, Pechlivanoglou P, Jalal HJ, Kao SYZ, Yang A, Enns EA. [A Need for Change! A Coding Framework for Improving Transparency in Decision Modeling](https://link.springer.com/article/10.1007/s40273-019-00837-x). [PharmacoEconomics](https://www.springer.com/journal/40273), 2190;37(11):1329–1339. https://doi.org/10.1007/s40273-019-00837-x

To run the CEA, you require [`dampack`: Decision-Analytic Modeling Package](https://cran.r-project.org/web/packages/dampack/index.html), an R package for analyzing and visualizing the health economic outputs of decision models.

## Use repository as a regular coding template
1. On the [tutorial's GitHub repository](https://github.com/DARTH-git/cohort-modeling-tutorial-intro), navigate to the main page of the repository (https://github.com/DARTH-git/cohort-modeling-tutorial-intro).
2. Above the file list, click **Clone or download** and select either 
    a. **Open in desktop**, which requires the user to have a GitHub desktop installed, or
    b. **Download zip** that will ask the user to download the whole repository as a .zip file.
3. Open the RStudio project `cohort-modeling-tutorial-intro.Rproj`.
4. Install all the required packages (as mentioned above)
    - [`dampack`](https://cran.r-project.org/web/packages/dampack/index.html)
    - [`darthtools`](https://github.com/DARTH-git/darthtools)
5. Run the scripts in the analysis folder.
6. Modify or adapt these scripts as needed for your project or analysis.

## Full list of Contributors:

  * [Fernando Alarid-Escudero](https://github.com/feralaes)
  
  * [Eline Krijkamp](https://github.com/krijkamp) 

  * [Eva Enns](https://github.com/evaenns)
  
  * [Alan Yang](https://github.com/alanyang0924)  
  
  * [Myriam Hunink](http://www.erasmus-epidemiology.nl/people/profile.php?id=45)
  
  * [Petros Pechlivanoglou](https://github.com/ppehli)
  
  * [Hawre Jalal](https://github.com/hjalal)
